---
title: "Hierarchical Bayesian Modeling of CO2 Flux (detailed diagnostics)"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Introduction

This document reproduces the `century_experiment.pdf` doc and focuses on the hierarchical models. I added the results obtained when solving the ODEs with a matrix exponential solution, and a few diagnostics to test for convergence.

## Century Model

### R code

The R function for simulating the data is shown below. The details are described above.
```{r simulate}
simulate_data_century <- function(t_meas, t_cap, init_C, num_rep) {
# INPUTS: 
#   t_meas:  measurement times
#   t_cap:   cap times
#   init_C:  initial pool contents
#   num_rep: number of replications
library(deSolve)
library(gtools)
genDerivs <-function(t, Ct, params) {
  # General diff eq model: dC_dt = I(t) + A(t)*C(t)
  # INPUTS:
  #   t: time
  #   Ct: the value of the vector C at time t, C(t)
  #   params: it has two fields, params$I and params$A
  dC_dt = params$I + params$A %*% Ct;
  return(list(dC_dt));
}
m <- 3; # number of pools
C_t0 <- matrix(init_C, nrow=m);
turnover <- c(1.5, 25, 1000);
K <- 1/turnover;
I <- rep(0, m); # no input flux for century
N_t <- length(t_meas);
CO2_flux_mat <- matrix(NA, nrow = N_t, ncol = num_rep);
Alpha_rep <- array(0, c(m, m, num_rep));
Alpha <- matrix(0, m, m);
# Setting global transfer rates with expert-tuned values:
Alpha[2, 1] = 0.5;
Alpha[3, 1] = 0.004;
Alpha[1, 2] = 0.42;
Alpha[3, 2] = 0.03;
Alpha[1, 3] = 0.45;
Alpha[1, 1] = 1 - Alpha[2, 1] - Alpha[3, 1];
Alpha[2, 2] = 1 - Alpha[1, 2] - Alpha[3, 2];
Alpha[3, 3] = 1 - Alpha[1, 3] - Alpha[2, 3];
for (this_rep in 1:num_rep) {
  # Hierarchical modeling of transfer rates for replications:
  kappa <- 10;
  Alpha_rep[,1, this_rep] <-  rdirichlet(1, Alpha[,1] * kappa);
  Alpha_rep[,2, this_rep] <-  rdirichlet(1, Alpha[,2] * kappa);
  Alpha_rep[,3, this_rep] <-  rdirichlet(1, Alpha[,3] * kappa);
  Alpha_rep[1, 1, this_rep] <- 0;
  Alpha_rep[2, 2, this_rep] <- 0;
  Alpha_rep[3, 3, this_rep] <- 0;
  A <- Alpha_rep[,, this_rep] * matrix(rep(K, m), nrow = m, byrow = TRUE) - diag(K);
  params <- list(I=I, A=A);
  t0 <- 0;
  # Solving the ODE system for given parameters:
  meas_data<-ode(y = C_t0, func = genDerivs,
          times = c(t0,t_meas), parms = params);
  cap_data<-ode(y = C_t0, func = genDerivs,
                 times = c(t0,t_cap), parms = params);
  # Calculating CO2 flux 
  totalC_t0 = sum(meas_data[1,2:(m+1)]);
  CO2_t_meas <- totalC_t0 - rowSums(meas_data[2:nrow(meas_data), 2:(m+1)]);
  CO2_t_cap <- totalC_t0 - rowSums(cap_data[2:nrow(cap_data),2:(m+1)]);
  CO2_flux <- (CO2_t_meas - CO2_t_cap)/(t_meas-t_cap); 
  # Adding log-normal noise:
  CO2_flux_mat[, this_rep] <- exp(log(CO2_flux) + rnorm(length(CO2_flux),0,.05));
}
simulated_data <- list(N_t = N_t, t_meas = t_meas, t_cap = t_cap, 
                       num_rep = num_rep, totalC_t0 = totalC_t0,
                       t0=t0, CO2_flux=CO2_flux_mat, Alpha_rep=Alpha_rep);
return(simulated_data);
}
```

## Model specification in Stan

We fit the simulated flux replications in the following way:
\begin{enumerate}
  \item Complete pooling: fitting a single model to all the replications, i.e., estimating only a single set of parameters for all models;
  \item No pooling: fitting a model separately to each replication, i.e., estimating the model parameters independently for each replication;
  \item Partial pooling: fitting a hierarchical Bayesian model which estimates parameters jointly for all replications and allows for variation between replicates. For this last model, the differential equation is solved in three manners:
  \begin{enumerate}
    \item using a stiff differential equation solver (bdf method)
    \item using a non-stiff differential equation solver (rk45 method)
    \item using a matrix exponential solution.
  \end{enumerate}
\end{enumerate}

### Data 

The data is also the same for all the models.

```
data { 
  real<lower=0> totalC_t0;      // initial total carbon 
  real t0;                      // initial time 
  int<lower=0> N_t;             // number of measurement times 
  int<lower=0> num_rep;         // number of replicates 
  vector<lower=t0>[N_t] t_meas; // measurement times 
  vector<lower=t0>[N_t] t_cap;  // cap times 
  matrix<lower=0>[N_t, num_rep] CO2_flux; // measured carbon fluxes 
} 
transformed data { 
  real x_r[0];  // no real data for ODE system 
  int x_i[0];   // no integer data for ODE system 
} 
```

Other parts of the stan model are different for different models. For a description of each model, see `century_experiments.pdf` and `century_experiments_me.htlm`. 


## Fitting the models
```{r fit, message=FALSE, warning=FALSE, cache=TRUE}
num_rep <- 5;
t_meas <- c(seq(from=1/360, to=7/360, by=1/360), seq(from=14/360, to=28/360, by=7/360),
            seq(from=60/360, to=360/360, by=30/360));
t_cap <- c(seq(from=0.5/360, to=6.5/360, by=1/360), seq(from=10.5/360, to=24.5/360, by=7/360),
           seq(from=45/360, to=345/360, by=30/360));
init_C <- c(.1, .1, .8);
data <- simulate_data_century(t_meas, t_cap, init_C, num_rep);

library(rstan);
library(tictoc);
rstan_options(auto_write = TRUE);
options(mc.cores = parallel::detectCores());
tic("hier_stiff")
fit_hier <- stan("century_hier.stan", data=data, iter=1500, seed=1234); # partial pooling
t3<-toc()
tic("hier_nonstiff")
fit_hier_nonstiff <- stan("century_hier_nonstiff.stan", data=data, iter=1500, seed=1234);
t4<-toc()
tic("hier_me")
fit_hier_me <- stan("century_hier_me.stan", data=data, iter=1500, seed=1234);
t5<-toc()
```


```{r, echo=FALSE, results="hide", message=FALSE, comment=NA}
library(rstan)
ss_hier<-summary(fit_hier)$summary
ss_hier_nonstiff<-summary(fit_hier_nonstiff)$summary
ss_hier_me<-summary(fit_hier_me)$summary
```

## initial carbon estimates ($\gamma$)

Actual Values: `gamma[1]=gamma[2]=0.1`, `gamma[3]=0.8`. 

### hierarchical (stiff solver)

```{r, echo=FALSE, message=FALSE, comment=NA}
print(ss_hier[c("gamma[1]", "gamma[2]", "gamma[3]"),], digits = 2)
```

### hierarchical (nonstiff solver)

```{r, echo=FALSE, message=FALSE, comment=NA}
print(ss_hier_nonstiff[c("gamma[1]", "gamma[2]", "gamma[3]"),], digits = 2)
```

### hierarchical (matrix exponential)

```{r, echo=FALSE, message=FALSE, comment=NA}
print(ss_hier_me[c("gamma[1]", "gamma[2]", "gamma[3]"),], digits = 2)
```

## turnover rates 

Actual Values: 1.5, 25, 1000.

### hierarchical (stiff solver)

```{r, echo=FALSE, message=FALSE, comment=NA}
print(ss_hier[c("turnover[1]", "turnover[2]", "turnover[3]"),], digits = 1)
```

### hierarchical (nonstiff solver)

```{r, echo=FALSE, message=FALSE, comment=NA}
print(ss_hier_nonstiff[c("turnover[1]", "turnover[2]", "turnover[3]"),], digits = 1)
```

### hierarchical (matrix exponential)

```{r, echo=FALSE, message=FALSE, comment=NA}
print(ss_hier_me[c("turnover[1]", "turnover[2]", "turnover[3]"),], digits = 1)
```


## transfer rates (a21)

Actual Global Value: 0.5 (with variations for each replication)

### pool

### hierarchical (stiff solver)

```{r, echo=FALSE, message=FALSE, comment=NA}
print(ss_hier[c("a21[1]", "a21[2]", "a21[3]", "a21[4]", "a21[5]"),], digits = 2)
```

### hierarchical (nonstiff solver)

```{r, echo=FALSE, message=FALSE, comment=NA}
print(ss_hier_nonstiff[c("a21[1]", "a21[2]", "a21[3]", "a21[4]", "a21[5]"),], digits = 2)
```


### hierarchical (matrix exponential)

```{r, echo=FALSE, message=FALSE, comment=NA}
print(ss_hier_me[c("a21[1]", "a21[2]", "a21[3]", "a21[4]", "a21[5]"),], digits = 2)
```

## transfer rates (a31)

Actual Global Value: 0.004 (with variations for each replication)

### pool

### hierarchical (stiff solver)

```{r, echo=FALSE, message=FALSE, comment=NA}
print(ss_hier[c("a31[1]", "a31[2]", "a31[3]", "a31[4]", "a31[5]"),], digits = 2)
```

### hierarchical (nonstiff solver)

```{r, echo=FALSE, message=FALSE, comment=NA}
print(ss_hier_nonstiff[c("a31[1]", "a31[2]", "a31[3]", "a31[4]", "a31[5]"),], digits = 2)
```

### hierarchical (matrix exponential)

```{r, echo=FALSE, message=FALSE, comment=NA}
print(ss_hier_me[c("a31[1]", "a31[2]", "a31[3]", "a31[4]", "a31[5]"),], digits = 2)
```


## transfer rates (a12)

Actual Global Value: 0.42 (with variations for each replication)

### hierarchical (stiff solver)

```{r, echo=FALSE, message=FALSE, comment=NA}
print(ss_hier[c("a12[1]", "a12[2]", "a12[3]", "a12[4]", "a12[5]"),], digits = 2)
```

### hierarchical (nonstiff solver)

```{r, echo=FALSE, message=FALSE, comment=NA}
print(ss_hier_nonstiff[c("a12[1]", "a12[2]", "a12[3]", "a12[4]", "a12[5]"),], digits = 2)
```

### hierarchical (matrix exponential)

```{r, echo=FALSE, message=FALSE, comment=NA}
print(ss_hier_me[c("a12[1]", "a12[2]", "a12[3]", "a12[4]", "a12[5]"),], digits = 2)
```


## transfer rates (a32)

Actual Global Value: 0.03 (with variations for each replication)

### pool

### hierarchical (stiff solver)

```{r, echo=FALSE, message=FALSE, comment=NA}
print(ss_hier[c("a32[1]", "a32[2]", "a32[3]", "a32[4]", "a32[5]"),], digits = 2)
```

### hierarchical (nonstiff solver)

```{r, echo=FALSE, message=FALSE, comment=NA}
print(ss_hier_nonstiff[c("a32[1]", "a32[2]", "a32[3]", "a32[4]", "a32[5]"),], digits = 2)
```

### hierarchical (matrix exponential)

```{r, echo=FALSE, message=FALSE, comment=NA}
print(ss_hier_me[c("a32[1]", "a32[2]", "a32[3]", "a32[4]", "a32[5]"),], digits = 2)
```


## transfer rates (a13)

Actual Global Value: 0.45 (with variations for each replication)

### hierarchical (stiff solver)

```{r, echo=FALSE, message=FALSE, comment=NA}
print(ss_hier[c("a13[1]", "a13[2]", "a13[3]", "a13[4]", "a13[5]"),], digits = 2)
```

### hierarchical (nonstiff solver)

```{r, echo=FALSE, message=FALSE, comment=NA}
print(ss_hier_nonstiff[c("a13[1]", "a13[2]", "a13[3]", "a13[4]", "a13[5]"),], digits = 2)
```

### hierarchical (matrix exponential)

```{r, echo=FALSE, message=FALSE, comment=NA}
print(ss_hier_me[c("a13[1]", "a13[2]", "a13[3]", "a13[4]", "a13[5]"),], digits = 2)
```


## Plotting fits to data

```{r plot_hier, echo=FALSE}
library(rstan)
library(ggplot2);
fit <- extract(fit_hier);
num_rep<-5;
CO2_flux_hat <- array(0,c(data$N_t, num_rep));
CO2_flux_lci <- array(NA,c(data$N_t, num_rep));
CO2_flux_uci <- array(NA,c(data$N_t, num_rep));
for (nr in 1:num_rep) {
  for (t in 1:data$N_t) {
    CO2_flux_hat[t,nr] <- median(fit$CO2_flux_hat[,t,nr]);
    CO2_flux_lci[t,nr] <- quantile(fit$CO2_flux_hat[,t,nr], 0.025);
    CO2_flux_uci[t,nr] <-  quantile(fit$CO2_flux_hat[,t,nr], 0.975);
  }
}
df_plot <- data.frame(list(t_meas = data$t_meas, 
                           CO2_flux_meas = data$CO2_flux,
                           CO2_flux_hat = CO2_flux_hat, CO2_flux_lci = CO2_flux_lci, 
                           CO2_flux_uci = CO2_flux_uci));

ggplot(df_plot, aes(x = t_meas)) +
  geom_ribbon(aes(ymin = CO2_flux_lci[,1],
                  ymax = CO2_flux_uci[,1]),
              fill="lightyellow") +
  geom_line(aes(y=CO2_flux_hat[,1]),colour="darkred") +
  geom_ribbon(aes(ymin = CO2_flux_lci[,2],
                  ymax = CO2_flux_uci[,2]),
              fill="lightyellow") +
  geom_line(aes(y=CO2_flux_hat[,2]),colour="darkred") +
  geom_ribbon(aes(ymin = CO2_flux_lci[,3],
                  ymax = CO2_flux_uci[,3]),
              fill="lightyellow") +
  geom_line(aes(y=CO2_flux_hat[,3]),colour="darkred") +
  geom_ribbon(aes(ymin = CO2_flux_lci[,4],
                  ymax = CO2_flux_uci[,4]),
              fill="lightyellow") +
  geom_line(aes(y=CO2_flux_hat[,4]),colour="darkred") +
  geom_ribbon(aes(ymin = CO2_flux_lci[,5],
                  ymax = CO2_flux_uci[,5]),
              fill="lightyellow") +
  geom_line(aes(y=CO2_flux_hat[,5]),colour="darkred") +
  geom_line(aes(y=data$CO2_flux[,1]),colour="darkblue") +
    geom_line(aes(y=data$CO2_flux[,2]),colour="darkblue") +
  geom_line(aes(y=data$CO2_flux[,3]),colour="darkblue") +
  geom_line(aes(y=data$CO2_flux[,4]),colour="darkblue") +
  geom_line(aes(y=data$CO2_flux[,5]),colour="darkblue") +
  labs(x="Time (days)", 
       y="CO2 flux") +
  ggtitle("Soil Incubation: Century Model (hierarchical-stiff),  \n data (blue), estimate median (red),  95% interval (yellow)");
```

```{r plot_hierns, echo=FALSE}
library(rstan)
library(ggplot2);
fit <- extract(fit_hier_nonstiff);
num_rep<-5;
CO2_flux_hat <- array(0,c(data$N_t, num_rep));
CO2_flux_lci <- array(NA,c(data$N_t, num_rep));
CO2_flux_uci <- array(NA,c(data$N_t, num_rep));
for (nr in 1:num_rep) {
  for (t in 1:data$N_t) {
    CO2_flux_hat[t,nr] <- median(fit$CO2_flux_hat[,t,nr]);
    CO2_flux_lci[t,nr] <- quantile(fit$CO2_flux_hat[,t,nr], 0.025);
    CO2_flux_uci[t,nr] <-  quantile(fit$CO2_flux_hat[,t,nr], 0.975);
  }
}
df_plot <- data.frame(list(t_meas = data$t_meas, 
                           CO2_flux_meas = data$CO2_flux,
                           CO2_flux_hat = CO2_flux_hat, CO2_flux_lci = CO2_flux_lci, 
                           CO2_flux_uci = CO2_flux_uci));

ggplot(df_plot, aes(x = t_meas)) +
  geom_ribbon(aes(ymin = CO2_flux_lci[,1],
                  ymax = CO2_flux_uci[,1]),
              fill="lightyellow") +
  geom_line(aes(y=CO2_flux_hat[,1]),colour="darkred") +
  geom_ribbon(aes(ymin = CO2_flux_lci[,2],
                  ymax = CO2_flux_uci[,2]),
              fill="lightyellow") +
  geom_line(aes(y=CO2_flux_hat[,2]),colour="darkred") +
  geom_ribbon(aes(ymin = CO2_flux_lci[,3],
                  ymax = CO2_flux_uci[,3]),
              fill="lightyellow") +
  geom_line(aes(y=CO2_flux_hat[,3]),colour="darkred") +
  geom_ribbon(aes(ymin = CO2_flux_lci[,4],
                  ymax = CO2_flux_uci[,4]),
              fill="lightyellow") +
  geom_line(aes(y=CO2_flux_hat[,4]),colour="darkred") +
  geom_ribbon(aes(ymin = CO2_flux_lci[,5],
                  ymax = CO2_flux_uci[,5]),
              fill="lightyellow") +
  geom_line(aes(y=CO2_flux_hat[,5]),colour="darkred") +
  geom_line(aes(y=data$CO2_flux[,1]),colour="darkblue") +
    geom_line(aes(y=data$CO2_flux[,2]),colour="darkblue") +
  geom_line(aes(y=data$CO2_flux[,3]),colour="darkblue") +
  geom_line(aes(y=data$CO2_flux[,4]),colour="darkblue") +
  geom_line(aes(y=data$CO2_flux[,5]),colour="darkblue") +
  labs(x="Time (days)", 
       y="CO2 flux") +
  ggtitle("Soil Incubation: Century Model (hierarchical-nonstiff),  \n data (blue), estimate median (red),  95% interval (yellow)");
```

```{r plot_hier, echo=FALSE}
library(rstan)
library(ggplot2);
fit <- extract(fit_hier_me);
num_rep<-5;
CO2_flux_hat <- array(0,c(data$N_t, num_rep));
CO2_flux_lci <- array(NA,c(data$N_t, num_rep));
CO2_flux_uci <- array(NA,c(data$N_t, num_rep));
for (nr in 1:num_rep) {
  for (t in 1:data$N_t) {
    CO2_flux_hat[t,nr] <- median(fit$CO2_flux_hat[,t,nr]);
    CO2_flux_lci[t,nr] <- quantile(fit$CO2_flux_hat[,t,nr], 0.025);
    CO2_flux_uci[t,nr] <-  quantile(fit$CO2_flux_hat[,t,nr], 0.975);
  }
}
df_plot <- data.frame(list(t_meas = data$t_meas, 
                           CO2_flux_meas = data$CO2_flux,
                           CO2_flux_hat = CO2_flux_hat, CO2_flux_lci = CO2_flux_lci, 
                           CO2_flux_uci = CO2_flux_uci));

ggplot(df_plot, aes(x = t_meas)) +
  geom_ribbon(aes(ymin = CO2_flux_lci[,1],
                  ymax = CO2_flux_uci[,1]),
              fill="lightyellow") +
  geom_line(aes(y=CO2_flux_hat[,1]),colour="darkred") +
  geom_ribbon(aes(ymin = CO2_flux_lci[,2],
                  ymax = CO2_flux_uci[,2]),
              fill="lightyellow") +
  geom_line(aes(y=CO2_flux_hat[,2]),colour="darkred") +
  geom_ribbon(aes(ymin = CO2_flux_lci[,3],
                  ymax = CO2_flux_uci[,3]),
              fill="lightyellow") +
  geom_line(aes(y=CO2_flux_hat[,3]),colour="darkred") +
  geom_ribbon(aes(ymin = CO2_flux_lci[,4],
                  ymax = CO2_flux_uci[,4]),
              fill="lightyellow") +
  geom_line(aes(y=CO2_flux_hat[,4]),colour="darkred") +
  geom_ribbon(aes(ymin = CO2_flux_lci[,5],
                  ymax = CO2_flux_uci[,5]),
              fill="lightyellow") +
  geom_line(aes(y=CO2_flux_hat[,5]),colour="darkred") +
  geom_line(aes(y=data$CO2_flux[,1]),colour="darkblue") +
    geom_line(aes(y=data$CO2_flux[,2]),colour="darkblue") +
  geom_line(aes(y=data$CO2_flux[,3]),colour="darkblue") +
  geom_line(aes(y=data$CO2_flux[,4]),colour="darkblue") +
  geom_line(aes(y=data$CO2_flux[,5]),colour="darkblue") +
  labs(x="Time (days)", 
       y="CO2 flux") +
  ggtitle("Soil Incubation: Century Model (hierarchical-stiff),  \n data (blue), estimate median (red),  95% interval (yellow)");
```
